Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-03-29 be.tex
conclusions,

We derive an expression for the bernoulli parameter of the gas, $\ke
+\gammaf \frac{p}{\rho}+\Phi$, assuming that the stellar density
profile $\rhostar\sim r^{\delta}$ inside of the break radius $\rb$.

@@

Where $\vw^2=v_w^2+\sigma^2= v_w^2+\frac{3}{\Gamma+2}\frac{G
  \Mbh}{r}+\sigma_0^2$.

Now integrate both sides with respect from r to the stagnation radius

@@

Where q is the rate of the mass injection per volume, which is
proportional to the stellar density. Let $x=r/\rs$ and $q_o= q
(\rs)$. Then $q=q_o x^{-\delta}$.

Then one obtains for the mass flux, $f=r^2 \rho v$,

@@

Recall that $\rs$ the stagnation radius, $x\equiv r/\rs$, $\zeta=\sqrt{v_w^2+\sigma_0^2}/\sigma_0$, $w\equiv (\rs/\rsoi)^{3\delta}$, and
$\delta$ is the power law slope of the stellar density profile inside
of the break radius, $\rb$.

% For sufficiently large radii ($x$), the expression above always
% becomes negative, which means that a break in the stellar density
% profile is necessary in order to obtain an outflow.
 For a given $\rb/\rsoi$ there is a critical $\zeta$, $\zeta_{c}$, such
 that if $\zeta>\zeta_c$, an outflow will be possible from any
radius, since the Bernoulli parameter will be positive inside of the
break radius. On the other hand if $\zeta<\zeta_c$, then there is a
minimum value of $(\rs/\rsoi)_{\rm min}$. If we have had
$\rs/\rsoi<(\rs/\rsoi)_{\rm min}$, then the Bernoulli parameter would
become negative inside of $\rb$ and no outflow would be possible.

For cusp galaxies ($\Gamma$=0.8), $\zeta_c\simeq \sqrt 2$
($v_w\simeq\sigma_0$).  On the other hand for core galaxies
($\Gamma$=0.1) we find $\zeta_c\simeq (\rb/\rsoi)^{0.45}$, see the top
panel of Fig.~\ref{fig:zetaCrit}. For values of $\zeta<\zeta_c$, the
stagnation radius rapidly approaches the break radius as shown in the
bottom panel of Fig.~\ref{fig:zetaCrit}.
% We also find that for core galaxies for $\zeta<\zeta_c$, there is a minimum
% for $\rs/\rsoi$, $\left(\rs/\rsoi\right)_{\rm min}$--if $\rs/\rsoi$ were
% less than this value the gas enthalpy would become negative inside of
% the break radius which is clearly unphysical.  This may be roughly
% approximated using (see Fig.~\ref{fig:rbRinf})

>>>>

We derive an expression for the bernoulli parameter of the gas, $\ke
+\gammaf \frac{p}{\rho}+\Phi$, assuming that the stellar density
profile $\rhostar\sim r^{-1-\Gamma}=r^{-\delta}$ inside of the break radius $\rb$

@@

Where $\vw^2=\sigma^2+v_w^2= \frac{3}{1+\delta}\frac{G
  \Mbh}{r}+\sigma_0^2+v_w^2$. The first term to the black hole
velocity dispersion, the second term corresponds to the velocity
dispersion of stars which we rake to be constant, and $v_w$
corresponds to any additional heating.

Now integrate both sides of equation~\eqref{eq:enCons} with respect from r to the stagnation radius $\rs$.

@@

Where q is the rate of the mass injection per volume, which is
proportional to the stellar density. Let $x=r/\rs$ and $q_o= q
(\rs)$. Then $q=q_o x^{-\delta}$. Thus,

@@

Recall that $\rs$ is the stagnation radius, $x\equiv r/\rs$, $\zeta=\sqrt{v_w^2+\sigma_0^2}/\sigma_0$, $w\equiv (\rs/\rsoi)^{3\delta}$, and
$\delta$ is the power law slope of the stellar density profile inside
of the break radius, $\rb$.

% For sufficiently large radii ($x$), the expression above always
% becomes negative, which means that a break in the stellar density
% profile is necessary in order to obtain an outflow.
 For a given $\rb/\rsoi$ there is a critical $\zeta$, $\zeta_{c}$, such
 that if $\zeta>\zeta_c$, an outflow will be possible from any
radius, since the Bernoulli parameter will be positive inside of the
break radius. On the other hand if $\zeta<\zeta_c$, then there is a
minimum value of $\rs/\rsoi$, $(\rs/\rsoi)_{\rm min}$. If we have had
$\rs/\rsoi<(\rs/\rsoi)_{\rm min}$, then the Bernoulli parameter would
become negative inside of $\rb$ and no outflow would be possible.

For cusp galaxies ($\Gamma$=0.8), $\zeta_c\simeq \sqrt 2$
($v_w\simeq\sigma_0$).  On the other hand for core galaxies
($\Gamma$=0.1, $\delta$=1.1) we find $\zeta_c\simeq
(\rb/\rsoi)^{0.44}$, see the top panel of Fig.~\ref{fig:zetaCrit}. For
values of $\zeta<\zeta_c$, the stagnation radius rapidly approaches
the break radius as shown in the bottom panel of
Fig.~\ref{fig:zetaCrit}.
% We also find that for core galaxies for $\zeta<\zeta_c$, there is a minimum
% for $\rs/\rsoi$, $\left(\rs/\rsoi\right)_{\rm min}$--if $\rs/\rsoi$ were
% less than this value the gas enthalpy would become negative inside of
% the break radius which is clearly unphysical.  This may be roughly
% approximated using (see Fig.~\ref{fig:rbRinf})

### 2015-03-29 ms.tex
conclusions,

\begin{table*}
\begin{threeparttable}
%\centering
\begin{minipage}{18cm}
\caption{Definitions of commonly used variables}
\begin{tabular}{ll}
\hline
{Variable} & {Definition} \\
\hline
$M_{\bullet}$ & Black hole mass \\
$\tilde{v}_{w}$ & Total heating parameter, including minimum heating
rate from stellar and black hole velocity dispersion \\
$v_{w}$ & Total heating parameter, excluding heating rate from velocity dispersion \\
$r_{s}$ & Stagnation radius (where gas radial velocity goes to zero) \\
$r_{\rm soi}$ & radius of sphere of influence (eq.~[\ref{eq:rsoi}]) \\
$r_{\rm b}$ & outer break radius of stellar density profile \\ 
$r_{\rm Ia}$ & radius interior to which Ia SN are infrequent compared to the dynamical/inflow timescale (eq.~[\ref{eq:rIa}]) \\ 
$\rho_{\star}(r)$ & 3D radial stellar density profile \\
$\rho(r)$ & gas density of CNM \\
$M_{\star}(r)$ & total enclosed stellar mass inside radius $r$ \\
$M_{\rm enc}(r)$ & total enclosed mass inside radius $r$ (SMBH + stars) \\
$q(r)$ & mass source term due to stellar winds (eq.~[\ref{eq:q}]) \\
$\eta$ & parameter setting normalization of mass input from stellar winds (eq.~[\ref{eq:q}]) \\
$\tau_{\star}$ & age of stellar population, in case of a single burst of star formation \\
$\Gamma$ & power-law slope of radial stellar density profile interior to the break radius \\
$\sigma(r)$ & stellar velocity dispersion \\
$\dot{M}$ & mass accretion rate \\
$\dot{M}_{\rm Ia}$ & maximum mass accretion rate as limited by SN Ia blow-out (eq.~[\ref{eq:eddr_Ia}]) \\
$\dot{M}_{\rm C}$ & equilibrium accretion rate set by Compton heating (eq.~[\ref{eq:MdotC}]) \\
$\dot{M}_{\rm TI}$ & maximum accretion rate for thermally stable accretion (eq.~[\ref{eq:Mdotmax}]) \\
$\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$ & ratio of external heating (stellar winds, SN Ia, MSPs) to radiative cooling (eq.~[\ref{eq:cooling2}]) \\
$t_{\rm h}$ & Hubble timescale \\
$\nu$ & gas density power-law slope at the stagnation radius (eq.~[\ref{eq:densSlope}]) \\
\hline
\label{table:definitions}  
\end{tabular}
\begin{tablenotes}
\item{$^{(a)}FILL IN$}
\end{tablenotes}
\end{minipage}
\end{threeparttable}

@@

  \begin{enumerate}
  %Younger population=>higher accretion rate; generally higher eta and
  %lower vw except for very young stellar populations.
  \item We find that for $\Mbh\gsim 10^6 \Msun$ and $\Mbh\lsim 10^8
    \Msun$, $\eddr \propto \Mbh^{\alpha}$ with $\alpha\lsim
    0.5 \, (0.8)$ for core (cusp) galaxies, 
    following average star formation histories measured in
    cosmological simulations \citep{MosterNaab+:2013a}. This is in
    tension observations of nuclear X-ray luminosity which suggest
    that $\eddr$ increases towards smaller $\Mbh$
    \citep{Miller+15}. However, it is important to note that the
    inflow rate on large scales (which we calculate) may differ from
    the actual accretion rate as outflows could occur on small
    scales. The outflow rate would depend on the angular momentum of
    the infalling gas and thus the angular momentum of stars in the
    nucleus. The in turn would vary systematically with $\Mbh$.
  \item We expect within scatter in nuclear x-ray luminosities at
    fixed $\Mbh$ from (1) stochastic variation in star formation
    history from galaxy to galaxy (2) Variations in small scale outflow
    rates due to variations between galactic nuclear angular
    momenta. {\bf AG: some more explanation of why 1 might not show up
    in Pellegrini...}
    These variations alone could explain much of the 2-4 orders of
    magnitude of scatter of nuclear x-ray luminosity at fixed $\Mbh$ 
    in the samples of \citet{Miller+15} and \citet{Pellegrini:2010a}.

>>>>

\begin{table*}
\begin{threeparttable}
%\centering
\begin{minipage}{18cm}
\caption{Definitions of commonly used variables}
\begin{tabular}{ll}
\hline
{Variable} & {Definition} \\
\hline
$M_{\bullet}$ & Black hole mass \\
$\tilde{v}_{w}$ & Total heating parameter, including minimum heating
rate from stellar and black hole velocity dispersion \\
$v_{w}$ & Total heating parameter, excluding heating rate from
velocity dispersion \\
$\sigma_0$ & Stellar velocity dispersion--assumed to be constant.\\
$\zeta$ & $\zeta=\sqrt{v_w^2+\sigma_0^2}/\sigma_0$\\
$r_{s}$ & Stagnation radius (where gas radial velocity goes to zero) \\
$r_{\rm soi}$ & radius of sphere of influence (eq.~[\ref{eq:rsoi}]) \\
$r_{\rm b}$ & outer break radius of stellar density profile \\ 
$r_{\rm Ia}$ & radius interior to which Ia SN are infrequent compared to the dynamical/inflow timescale (eq.~[\ref{eq:rIa}]) \\ 
$\rho_{\star}(r)$ & 3D radial stellar density profile \\
$\rho(r)$ & gas density of CNM \\
$M_{\star}(r)$ & total enclosed stellar mass inside radius $r$ \\
$M_{\rm enc}(r)$ & total enclosed mass inside radius $r$ (SMBH + stars) \\
$q(r)$ & mass source term due to stellar winds (eq.~[\ref{eq:q}]) \\
$\eta$ & parameter setting normalization of mass input from stellar winds (eq.~[\ref{eq:q}]) \\
$\tau_{\star}$ & age of stellar population, in case of a single burst of star formation \\
$\Gamma$ & power-law slope of radial stellar surface brightness
profile interior to the break radius \\
$\delta$ & power-law slope of the 3d stellar density profile inside of
the break radius. $\delta=\Gamma+1$.\\
$\dot{M}$ & mass accretion rate \\
$\dot{M}_{\rm Ia}$ & maximum mass accretion rate as limited by SN Ia blow-out (eq.~[\ref{eq:eddr_Ia}]) \\
$\dot{M}_{\rm C}$ & equilibrium accretion rate set by Compton heating (eq.~[\ref{eq:MdotC}]) \\
$\dot{M}_{\rm TI}$ & maximum accretion rate for thermally stable accretion (eq.~[\ref{eq:Mdotmax}]) \\
$\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$ & ratio of external heating (stellar winds, SN Ia, MSPs) to radiative cooling (eq.~[\ref{eq:cooling2}]) \\
$t_{\rm h}$ & Hubble timescale \\
$\nu$ & gas density power-law slope at the stagnation radius
(eq.~[\ref{eq:densSlope}]) \\
\hline
\label{table:definitions}  
\end{tabular}
\begin{tablenotes}
\item{$^{(a)}FILL IN$}
\end{tablenotes}
\end{minipage}
\end{threeparttable}

@@

  \begin{enumerate}
  %Younger population=>higher accretion rate; generally higher eta and
  %lower vw except for very young stellar populations.
  \item The gas will be divided between an accretion flow and an
    outward flow at the stagnation radius, $\rs$. In a steady state
    the gas inflow rate towards the hole will be proportional to the
    stellar mass enclosed inside of $\rs$. 
  \item The location of the stagnation radius and thus the accretion
    rate is quite sensitive to the effective heating rate of the
    gas. In our model gas is supplied by stellar winds and the kinetic
    energy of the winds is thermalized as the winds
    collide. Supernovae and AGN feedback may provide additional
    sources of heating. Each source of heating is related to the star
    formation history of the galaxy.  For example, a galaxy with star
    formation within the last $\sim10^7$ years would have mass
    injection from very fast stellar winds ($w_w\simeq 1000$ km/s). In
    contrast, an older population would have heating only from AGN
    feedback and supernovae.

    Quantitatively $\rs\sim v_w^{-2}$, which implies $\Mdot\sim
    v_w^{-3.8 (-2.4)}$ for core (cusp galaxies). However, $\Mdot$ may
    increase more rapidly for small $v_w$ for two reasons. (1) When
    $v_w$ becomes comparable to the stellar velocity dispersion, the
    gas inside of the break radius inside of the break radius inside
    of the stellar density profile is bound and cannot move
    outflow. (2) For $v_w\simeq200 $ km/s the gas may become thermally
    unstable and there may be a cycle of alternating high and low
    accretion rates onto the BH as described in...{\bf AG: Fill in}

  \item We find that for $\Mbh\gsim 10^6 \Msun$ and $\Mbh\lsim 10^8
    \Msun$, $\eddr \propto \Mbh^{\alpha}$ with $\alpha\lsim
    0.5 \, (0.8)$ for core (cusp) galaxies, 
    following average star formation histories measured in
    cosmological simulations \citep{MosterNaab+:2013a}. This is in
    tension observations of nuclear X-ray luminosity which suggest
    that $\eddr$ increases towards smaller $\Mbh$
    \citep{Miller+15}. However, it is important to note that the
    inflow rate on large scales (which we calculate) may differ from
    the actual accretion rate as outflows could occur on small
    scales. The outflow rate would depend on the angular momentum of
    the infalling gas and thus the angular momentum of stars in the
    nucleus. The in turn would vary systematically with $\Mbh$.
  \item We expect scatter in nuclear x-ray luminosities at
    fixed $\Mbh$ from (1) stochastic variation in star formation
    history from galaxy to galaxy (2) Variations in small scale outflow
    rates due to variations between galactic nuclear angular
    momenta. {\bf AG: some more explanation of why 1 might not show up
    in Pellegrini...}
    These variations alone could explain much of the 2-4 orders of
    magnitude of scatter of nuclear x-ray luminosity at fixed $\Mbh$ 
    in the samples of \citet{Miller+15} and \citet{Pellegrini:2010a}.
